### 2021-10-08 docs/paper/sections/0_abstract.tex
Revise paper and apply feedback by @scheibel
- Restructure sections
- Shorten things
- Rephrasing
- Improve typesetting

	In the dependency graph of a software ecosystem, \emph{downstream dependencies} are the nodes that depend on a package.
	Other than for upstream dependencies, solutions that provide individual package developers with this kind of information rarely exist to date.
	This paper makes two major contributions:
	\begin{enumerate*}[label=(\roman*)]
		\item We propose an approach for mining usage data about packages from their downstream dependencies by utilizing a static type analyzer.
		\item We introduce a tool that allows npm package developers to survey usage samples of their APIs by incorporating the aggregated usage data into an integrated development environment.%
			\footnote{
				All artifacts are available on GitHub under the following URL: \url{https://github.com/LinqLover/downstream-repository-mining}
				\TODO{Insert name and use link to VSCE}
			}
	\end{enumerate*}
	We find that usage data from downstream dependency repositories are a promising source of information for mining software repositories and that they can be useful to support package developers in improving or extending their APIs.

>>>>

	In the dependency graph of a software ecosystem, \emph{downstream dependencies} are the nodes that depend on a package.
	Other than for upstream dependencies, solutions that provide individual package developers with this kind of information rarely exist to date.
	This paper makes two major contributions:
	\begin{enumerate*}[label=(\roman*)]
		\item We propose an approach for gathering downstream dependencies of a single package efficiently and extracting usage samples from them by utilizing a static type analyzer.
		\item We introduce a tool that allows npm package developers to survey usage samples of their APIs by incorporating the aggregated usage data into an IDE.%
			\footnote{
				All artifacts are available on GitHub under the following URL: \url{https://github.com/LinqLover/downstream-repository-mining}
				\TODO{Insert name and use link to VSCE}
			}
	\end{enumerate*}

	Our two methods yield over \num{8000} dependencies for popular packages at an average precision of \SI{98.7}{\percent} or \SI{68.4}{\percent}, resp., but tend to exclude unpopular dependencies.
	The extracted usage samples have a high precision but are unavailable for repositories with complex build configurations or metaprogramming patterns.
	Our tool is able to process up to 12 dependencies per minute while requiring about \SI{500}{\mega\byte} RAM in total and less than \SI{30}{\mega\byte} ROM per package.
	We find that usage data from downstream dependency repositories are a promising source of information for mining software repositories and that our approach could support package developers in improving or extending their APIs.

### 2021-10-08 docs/paper/sections/1_introduction.tex
Revise paper and apply feedback by @scheibel
- Restructure sections
- Shorten things
- Rephrasing
- Improve typesetting

Open Source software (OSS) solutions have become more and more and important during the last decades.
Especially, this trend has been experiencing an additional updraft due to the ongoing spreading of OSS platforms such as GitHub or GitLab.
Open-source development offers many advantages as opposed to traditional closed-source development, including large numbers of volunteer contributions from the open-source community, increased transparency effects in security-related domains, and a high potential for reusing solutions.
These solutions are usually organized as \emph{packages} each of which attempts to solve an isolated problem and provide a generic interface.
Most commonly, packages are developed in a \emph{code repository} that is managed using a \emph{development platform} (such a \emph{GitHub}\footnote{\url{https://github.com/}}, \emph{GitLab}\footnote{\url{https://gitlab.com/}}, or \emph{Bitbucket}\footnote{\url{https://bitbucket.com/}}) and deployed to a \emph{package repository} using a \emph{package manager} (such as \emph{PyPI}\footnote{\url{https://pypi.org/}} using \emph{pip}\footnote{\url{https://pip.pypa.io/}} for the Python programming language, the \emph{npm registry}\footnote{\url{https://www.npmjs.com/}} for JavaScript/TypeScript, or the \emph{NuGet Gallery}\footnote{\url{https://www.nuget.org/}} for languages from the .NET ecosystem).
Other solutions or packages can then \emph{depend} on existing packages by declaring them in their manifest, i.e., the \code{package.json} file of an npm package, or the \code{requirements.txt} file of a Python project.
These dependency relations form a large directed graph that connects major parts of the software world for each popular programming language ecosystem.

Despite this connectedness by design, however, the development process of many packages is still characterized by an isolated approach:
While OSS developers commonly submit tickets to or contribute patches against \emph{upstream repositories} that they depend on to solve subproblems, the reverse direction of these edges -- called \emph{downstream dependencies} -- is often neglected by package developers when they extend, restructure, or refactor their solutions.
This knowledge deficit can cause a wide range of alignment issues, including poorly suited interfaces \citep{piccioni2013empirical}, unidentified defects \citep{wong2017more}, and compatibility problems \citep{bogart2015breaks}.
Eventually, all these issues impair the capabilities of the global OSS community to build and support high-quality products.
In the following, we also refer to downstream dependencies will shortly as \emph{dependencies}.

@@

The rest of this paper is organized as follows:
In \cref{sec:related_work}, we give a summary of existing approaches in this field.
In \cref{sec:framework}, we outline the overall conditions and the general approach for our solution to mining downstream dependencies.
In \cref{sec:dependency_collection} and \cref{sec:usage_mining}, we explain our approach for collecting downstream dependency repositories and mining usage data from these repositories in detail, resp.
In \cref{sec:implementation}, we describe an implementation of the proposed approach and our design decisions for displaying usage samples to the user.
In \cref{sec:evaluation}, we examine the fitness of our data mining approach and the usability of the display solution.

>>>>

Open Source software (OSS) solutions have become more and more and important during the last decades.
Especially, this trend has been experiencing an additional updraft due to the ongoing spreading of OSS platforms such as GitHub or GitLab.
Open-source development offers many advantages as opposed to traditional closed-source development, including volunteer contributions from the open-source community, increased transparency effects in security-related domains, and a high potential for reusing solutions.
These solutions are usually organized as \emph{packages} each of which attempts to solve an isolated problem and provide a generalized interface.
Most commonly, packages are developed in a \emph{code repository} that is managed using a \emph{development platform} (such a \emph{GitHub}\footnote{\url{https://github.com/}}, \emph{GitLab}\footnote{\url{https://gitlab.com/}}, or \emph{Bitbucket}\footnote{\url{https://bitbucket.com/}}) and deployed to a \emph{package repository} using a \emph{package manager} (such as \emph{PyPI}\footnote{\url{https://pypi.org/}} using \emph{pip}\footnote{\url{https://pip.pypa.io/}} for the Python programming language, the \emph{npm registry}\footnote{\url{https://www.npmjs.com/}} for JavaScript, or the \emph{NuGet Gallery}\footnote{\url{https://www.nuget.org/}} for .NET languages).
Other solutions or packages can then \emph{depend} on existing packages by declaring them in their manifest, i.e., the \code{package.json} file of an npm package, or the \code{requirements.txt} file of a Python project.
These dependency relations form a large directed graph that connects major parts of the software world for each popular programming language ecosystem.

Despite this connectedness by design, however, the development process of many packages is still characterized by an isolated approach:
While OSS developers commonly submit tickets to or contribute patches against \emph{upstream repositories} that they depend on to solve subproblems, the reverse direction of these edges -- called \emph{downstream dependencies} -- is often neglected by package developers when they extend, restructure, or refactor their solutions.
This knowledge deficit can cause a wide range of alignment issues, including poorly suited interfaces \citep{piccioni2013empirical}, unidentified defects \citep{wong2017more}, and compatibility problems \citep{bogart2015breaks}.
Eventually, all these issues impair the capabilities of the global OSS community to build and support high-quality products.
In the following, we also refer to downstream dependencies shortly as \emph{dependencies}.

@@

In essence, our proposed approach consists of three steps:
\begin{enumerate*}[label=(\roman*)]
	\item collect downstream dependency repositories from public clouds,
	\item mine package usage samples from these dependencies by using a type analyzer,
	and \item aggregate and present these usage data to the user.
\end{enumerate*}

The rest of this paper is organized as follows:
In \cref{sec:related_work}, we give a summary of existing approaches in this field.
In \cref{sec:approach}, we outline the overall conditions for our solution and explain our approach for collecting downstream dependency repositories and mining usage data from these repositories.
In \cref{sec:implementation}, we describe an implementation of the proposed approach and our design decisions for displaying usage samples to the user.
In \cref{sec:evaluation}, we examine the fitness of our data mining approach and the usability of the display solution.

### 2021-10-08 docs/paper/sections/2_related_work.tex
Revise paper and apply feedback by @scheibel
- Restructure sections
- Shorten things
- Rephrasing
- Improve typesetting

\section{Related work}
\label{sec:related_work}

@@

\subsection{Searching dependency graphs}
\label{sec:related_work/dependencies}

@@

However, existing research about dependency graphs primarily takes a broad statistical perspective, and researchers often rely on a large corpus of downloaded software repositories \citep{abdalkareem2017developers,katz2020libraries,kikas2017structure}.
As opposed to this approach, these capacities will not be appropriate for scenarios that are occasionally performed by developers of a single package only.
In this case, developers will often use a public \emph{code search} service such as GitHub code search\footnote{\url{https://docs.github.com/en/github/searching-for-information-on-github}} or Sourcegraph\footnote{\url{https://sourcegraph.com/}} instead.
\citet{liu2020opportunities} provide a (not yet published) survey of current methods and trends in code search tools that include, besides simple full-text search, newer approaches such as structural or semantic search queries, code similarity metrics, or Machine Learning methods.

\subsection{Extracting usage samples}
\label{sec:related_work/usage_samples}

@@

\citet{collard2013srcml} propose an infrastructure called \caps{srcML} that aims to create a unified representation of multilingual source code snippets for arbitrary analysis purposes, including the construction of call graphs.
Another solution is proposed by \citet{bogar2018lightweight} who focus on analyzing multilingual codebases and who utilize an island parser to build call graphs.
Furthermore, \citet{antal2018static} compare several call graph generators for JavaScript.
They also emphasize that for languages supporting dynamic typing or meta-programming mechanisms, static call graph generators have a limited precision by design.

To bring together dependency graphs and call graphs, \emph{ecosystem call graphs} can be constructed by applying call graphs to entire ecosystems, crossing repository boundaries.
\citet{hejderup2018software} propose an approach to do so for the JavaScript/npm ecosystem in which they also model the semantic version numbers of packages and conduct an impact analysis for vulnerabilities based on their graph; a very similar method is described by \citet{boldi2020fine} to formulate the vision of a new design for package management systems that provides options for finer-grained reuse of solutions.
\citet{nielsen2021modular} elaborate on this work by developing a call graph generator tailored to this problem that comes with significant improvements in terms of recall and performance.
\citet{wang2020empirical} describe a similar approach for the same ecosystem and apply it to the domain of locating security issues in a project's upstream dependencies.
With \caps{Pr\"azi} (not yet peer-reviewed), \citet{hejderup2021praezi} also describe their implementation of a dependency-scale call graph for the Rust/Cratesio ecosystem in-depth.
For Java/Maven, \citet{keshani2021scalable} proposes another solution.
Another approach is implemented in the \emph{precise code intelligence} feature of the Sourcegraph code search engine that makes code references explorable across repository boundaries for repositories that provide a metadata file in the \emph{Language Server Index Format}.\footnote{\url{https://docs.sourcegraph.com/code_intelligence/explanations/precise_code_intelligence}}

@@

\subsection{Presentation of results}
\label{sec:related_work/presentation}

To establish an adequate developer experience, any collected usage data still need to be presented suitably.
A common approach for this is a hierarchical and navigatable representation of the collected call tree.
An early form of this representation has been invented as the interactive ``message set'' tool for browsing senders and implementors in an object-oriented system for the Smalltalk programming environment \citep[section~10.1f.]{goldberg1984smalltalk}.
Modern alternatives include the \caps{Stacksplorer} \citep{karrer2011stacksplorer} or \caps{Blaze} \citep{kramer2012blaze} that list next to the currently focused method other methods which are adjacent to this method in the call graph.
With a focus on exploring references to API members, \citet{de2013multi} propose a set of additional views, including hierarchical lists and word clouds for highly referenced identifiers.
Focusing on exploring APIs in a non-code-centric way, \citet{hora2015apiwave} propose a dashboard that allows developers to browse existing APIs and track possibly breaking changes in their interfaces.

>>>>

\section{Related Work}
\label{sec:related_work}

@@

\paragraph{Searching dependency graphs}
\label{sec:related_work/dependencies}

@@

However, existing research about dependency graphs primarily takes a broad statistical perspective, and researchers often rely on a large corpus of downloaded software repositories \citep{abdalkareem2017developers,katz2020libraries,kikas2017structure}.
As opposed to this approach, these capacities will not be appropriate for scenarios occasionally performed by developers of a single package only.
In this case, developers will often use a public \emph{code search} service such as \emph{GitHub} code search\footnote{\url{https://docs.github.com/en/github/searching-for-information-on-github}} or \emph{Sourcegraph}\footnote{\url{https://sourcegraph.com/}} instead.
\citet{liu2020opportunities} provide a pre-printed survey of current methods and trends in code search tools that include, besides simple text search, newer approaches such as structural or semantic search queries, code similarity metrics, or Machine Learning methods.

\paragraph{Extracting usage samples}
\label{sec:related_work/usage_samples}

@@

\citet{collard2013srcml} propose an infrastructure called \caps{srcML} that aims to create a unified representation of multilingual source code snippets for arbitrary analysis purposes, including the construction of call graphs.
Another solution is proposed by \citet{bogar2018lightweight} who utilize an island parser to build polyglot call graphs.
Furthermore, \citet{antal2018static} compare several call graph generators for JavaScript.
They also emphasize that for languages supporting dynamic typing or meta-programming mechanisms, static call graph generators have a limited precision by design.

To bring together dependency graphs and call graphs, \emph{ecosystem call graphs} can be constructed by applying call graphs to entire ecosystems, crossing repository boundaries.
Many approaches apply this concept to different ecosystems, often aiming to conduct a fine-grained impact analysis for security vulnerabilities \citep{boldi2020fine,hejderup2018software,hejderup2021praezi,keshani2021scalable,nielsen2021modular,wang2020empirical}.
Another approach is implemented in the \emph{precise code intelligence} feature of the Sourcegraph code search engine that makes code references explorable across repository boundaries for repositories that provide a metadata file in the \emph{Language Server Index Format}.\footnote{\url{https://docs.sourcegraph.com/code_intelligence/explanations/precise_code_intelligence}}

@@

\paragraph{Presentation of results}
\label{sec:related_work/presentation}

To establish an adequate developer experience, any collected usage data still need to be presented suitably.
A common approach for this is a hierarchical and navigatable representation of the collected call tree.
An early form of this representation has been invented as the interactive ``message set'' tool for browsing senders and implementors in an object-oriented system for the Smalltalk programming environment \citep{goldberg1984smalltalk}.
Modern alternatives include the \emph{Stacksplorer} \citep{karrer2011stacksplorer} or \emph{Blaze} \citep{kramer2012blaze} that list next to the currently focused method other methods which are adjacent to this method in the call graph.
With a focus on exploring references to API members, \citet{de2013multi} propose a set of additional views, including hierarchical lists and word clouds for highly referenced identifiers.
Focusing on exploring APIs in a non-code-centric way, \citet{hora2015apiwave} propose a dashboard that allows developers to browse existing APIs and track possibly breaking changes in their interfaces.

### 2021-10-08 docs/paper/sections/3_approach.tex
Rework first sections (up to approach)
Revise paper and apply feedback by @scheibel
- Restructure sections
- Shorten things
- Rephrasing
- Improve typesetting

\section{Approach}
\label{sec:approach}

Because of its superior popularity \TODO{prove}, in this work, we will confine ourselves to the JavaScript/npm ecosystem.
Our goal is thus to develop a tool for npm package developers that displays an overview of the downstream dependencies and their details for a chosen package.
Concretely, we pose the following key requirements to this tool:

\begin{enumerate}[label=R\arabic*]
	\item \label{req1} The tool allows package developers to survey and explore the usages of individual package members.
	\item The tool blends in well with the usual workflow of package developers.
	\item \label{req3} The tool runs out-of-the-box with a small footprint, without the user needing to perform any sophisticated setup steps or providing additional system resources for it to work.
\end{enumerate}

In essence, our proposed approach to fulfill these requirements consists of three steps:
\begin{enumerate*}
	\item collect downstream dependency repositories,
	\item mine package usage samples from them,
	and \item aggregate and present these usage data to the user.
\end{enumerate*}
In the following, we will describe each step.

\subsection{Downstream dependency collection}
\label{sec:approach/collection}

In the first step, a list of repositories that depend on the target package has to be assembled.
As mentioned in \cref{sec:related_work/usage_information}, many approaches start with a large downloaded corpus of unspecific repositories from the ecosystem which they then iterate over to filter the relevant repositories.
However, a major drawback of this approach is the high resource demand for both creating and traversing this corpus.
This approach is suited for analyzing repositories in large scale but is not conform to \cref{req1} due to the large footprint in terms of computational power and time.

As an alternative, we have decided to apply significantly stricter pre-filtering to the set of repositories before downloading it to the local machine.
For that, we have chosen two publicly available data sources:\footnote{
	Other data sources that we have rejected include GitHub code search which has limited precision results and is not freely accessible programatically as well as the \emph{libraries.io} service \citep{katz2020libraries} whose downstream dependency search was inoperative at the time of writing.%
}

\begin{enumerate}[label=(\roman*)]
	\item The npm registry which maintains a doubly-connected edge list of all npm packages that depend on each other.
	\item The code search engine \emph{Sourcegraph}\footnote{\url{https://sourcegraph.com/}} which is indexing repositories from most popular OSS platforms such as GitHub or GitLab.
		Using this search engine, we can query all repositories that declare a dependency on the target package in their package metadata file.
\end{enumerate}

Finally, we can download the source code for a small number of relevant repositories only.

\subsection{Usage sample mining}
\label{sec:approach/mining}

\begin{figure*}[htb!]
	\newcommand\lowlight[1]{\textcolor{gray}{#1}}
	\newcommand\highlight[1]{\textcolor{red}{#1}}
	\centering
	\begin{subfigure}[t]{.32\linewidth}
		\include{sections/3_approach_assets/patterns/CallExpression}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{.32\linewidth}
		\include{sections/3_approach_assets/patterns/PropertyAccessExpression}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{.32\linewidth}
		\include{sections/3_approach_assets/patterns/JsxOpeningExpression}
	\end{subfigure}

	\caption{Example AST patterns for TypeScript programs.
		The \highlight{highlighted} node contains the link to the declaration of the referenced identifier.
	}
	\label{fig:approach/mining/patterns}
\end{figure*}

Having downloaded the selected downstream dependency repositories, we can proceed to extract usage samples for the target package from each dependency repository.
Our goal is to identify these usage samples on a fine-granular level so that we can trace them down to the single identifiers that were exposed by the target package.

To do so, we start by parsing the source code of every dependency repository as well as the source of the target package each into a separate AST.

% TODO: use term "binding"?
In a second step, every dependency AST is traversed together with the target package AST in order to create links between semantically related nodes, e.g., between an expression and the variable it is assigned to, or between a function call and the definition of this function.

In the final step, we traverse the AST again and collect all nodes that are linked to a declaration in the target package.
To identify these links, we define a number of patterns for node structures and look up the declaration of every matched node (see \cref{fig:approach/mining/patterns}).

\subsection{Presentation of results}
\label{sec:approach/presentation}

After all references have been collected, a proper UI is still required to provide users easy access to these data that fulfills the requirements mentioned above.
To satisfy \cref{req2} and make our tool avaiable in the usual working environment of users, we implement it as an extension to a popular IDE.
One of the most popular IDEs that support JavaScript, and the one with the highest annual growth, is Visual Studio Code\footnote{\url{https://code.visualstudio.com/}}\footnote{Top IDE Index by Pierre Carbonnelle (2021): \url{https://pypl.github.io/IDE.html}}.
As it also provides a comprehensive set of APIs for extension developers, we have decided to roll our tool out as a VS Code extension.

For choosing appropriate views of the raw reference data, we have identified three common questions of package developers to answer each of which they require access to downstream dependencies and their usage samples:

\begin{enumerate}[label=Q\arabic*]
	\item Which dependencies do use the target package and what problems are they solving with it?
	\item By how many dependencies is a particular package member being used?
	\item In which contexts and constellations is a particular package member being used?
\end{enumerate}

To support developers in answering these questions and fulfill \cref{req1}, we have designed the following views for our extension:

\begin{enumerate}[label=(\roman*)]
	\item The \emph{dependency browser} allows to explore all downstream dependencies and, grouped for each dependency, all their references to the target package.
	\item The \emph{usage browser} displays all public package member and, grouped for each member, all dependencies and their references to this member.
	\item The \emph{CodeLens integration} provides quick access to a slice of the usage browser that is attached to the definition of each package member right in the source code editor.
\end{enumerate}

Both references and members are organized in a tree view reflecting the hierarchical structure of the original software repository.
In addition, we recognize the effort of browsing large lists of dependency data and encounter it by providing an ``I'm felling lucky'' button for every view that redirects the user to a random dependency resp. reference to gain a faster, unbiased impression of usage samples.

>>>>

\section{Downstream Dependency Mining}
\label{sec:approach}

To improve the developers' knowledge about the usages of their packages, our goal is to build a system that allows package developers to survey the API usage of downstream dependencies for a chosen target package.
Concretely, we have identified three common questions of package developers about the usage of their packages:

\begin{enumerate}[label=Q\arabic*]
	\item \label[question]{q1} Which dependencies are using the target package, and what problems do they attempt to solve with it?
	\item \label[question]{q2} By how many dependencies is a particular package member being used?
	\item \label[question]{q3} In which contexts and constellations is a particular package member being used?
\end{enumerate}

Furthermore, we pose two organizational requirements to the eventual tool solution to ensure its practical applicability for the target user group:

\begin{enumerate}[label=R\arabic*]
	\item \label[requirement]{req1} The tool runs out-of-the-box with a small footprint, without the user needing to perform any sophisticated setup steps or providing additional system resources.
	\item \label[requirement]{req2} The tool blends in with the usual workflow of package developers at the best possible rate.
\end{enumerate}

In the following, we will describe both steps of our approach to collect the desired data.

\subsection{Downstream Dependency Collection}
\label{sec:approach/dependency_collection}

In the first step, a list of repositories that depend on the target package has to be assembled.
As mentioned in \cref{sec:related_work/usage_samples}, many approaches start with a large downloaded corpus of unspecific repositories from the ecosystem, which they then iterate over to filter the relevant repositories.
However, a major drawback of this approach is the high resource demand for both creating and traversing this corpus.
This approach is suited for analyzing repositories on a large scale but is not conform to \cref{req1} because of its large footprint in terms of computational power and time.

As an alternative, we have decided to apply significant pre-filtering to the set of repositories before downloading a subset of it to the local machine.
For that, we have identified two types of data sources that are available in public clouds:

\begin{enumerate}[label=(\roman*)]
	\item Package repositories that maintain a doubly-connected edge list of interdependent packages in an ecosystem.
	\item Code search engines that index the source code of many repositories from OSS platforms.
		Using these search engines, we can query all repositories that declare a dependency on the target package in their package manifest file.
\end{enumerate}

\subsection{Usage Sample Mining}
\label{sec:approach/usage_mining}

\begin{figure*}
	\begin{minipage}{\linewidth}
		\newcommand\textlowlight[1]{\textcolor{gray}{#1}}
		\newcommand\texthighlight[1]{\textcolor{accent1}{#1}}
		\begin{center}
			\begin{subfigure}[t]{.32\linewidth}
				\input{sections/3_approach/usage_mining/patterns/CallExpression}
			\end{subfigure}
			\hfill
			\begin{subfigure}[t]{.32\linewidth}
				\input{sections/3_approach/usage_mining/patterns/PropertyAccessExpression}
			\end{subfigure}
			\hfill
			\begin{subfigure}[t]{.32\linewidth}
				\input{sections/3_approach/usage_mining/patterns/JsxOpeningElement}
			\end{subfigure}

			\caption{AST patterns for example JavaScript/TypeScript expressions.
				The \texthighlight{highlighted} node contains the link to the declaration of the referenced identifier.
			}
			\label{fig:approach/usage_mining/patterns}
		\end{center}
	\end{minipage}
\end{figure*}

Having downloaded the selected downstream dependency repositories, we can proceed to extract usage samples for the target package from each dependency repository.
Our goal is to collect fine-grained source code excerpts that reference single identifiers exposed by the target package.
The complete procedure is displayed in \cref{alg:usage_mining}.

\begin{algorithm}[b]
	\caption{Extraction of usage samples.}\label{alg:usage_mining}

	\KwIn{\\\Indp
		$\mli{pkg}$: target package \\
		$\mli{dependencies}$: downstream dependencies}
	\KwOut{usage samples (set of strings)}
	\;
	\For{$\mli{dep} \in \mli{dependencies}$}{
		$\mli{asf} \gets \mtx{parse}(\mli{dep} \cup \mli{pkg})$\;
	  	$\mtx{annotate\_types}(\mli{asf})$\;
		\For{$\mli{ast} \in \mli{asf}$}{
			\For{$\mli{node} \in \mli{dfs}(\mli{ast})$}{
				\For{$\mli{pattern} \in \mli{patterns}$}{
					\If{$\mli{pattern}.\mtx{matches}(\mli{node}) \wedge \mli{pkg}.\mtx{declares}(\mli{pattern}.\mtx{getType}(\mli{node}))$}{
						\kwYield{$\mli{node}.\mtx{text}$}
					}
				}
			}
		}
	}
\end{algorithm}

To do so, we start by parsing the source code of every dependency repository as well as the source code of the target package each into a separate abstract syntax forest (ASF, i.e., a set of ASTs).

In a second step, we perform a static type analysis against each dependency ASF together with the target package's ASF.
The results of the type analysis are attached to the ASF so that every identifier is annotated with a type symbol that links to the declaration of the type for the identifier.
For instance, after this step, every variable node will contain a link to the assignment node of this variable, and every function call expression will contain a link to the definition of this function.
The operating principle of this type analysis depends on the kind of programming language; i.e., in statically typed languages, type symbols can usually be retrieved from the declaration of an identifier, whereas in dynamically typed languages, a control flow analysis will be required to identify the origin of every identifier's type.

In the final step, from each ASF, we collect all nodes whose type is declared in the target package.
To identify these links, we define a set of language-specific patterns for AST subtrees that constitute a usage expression.
In particular, each of these patterns expects a node containing a link to an identifier declaration (see \cref{fig:approach/usage_mining/patterns}).
